Bayesian inference of relative fitness on high-throughput pooled competition assays

### 2023-12-05 doc/_experiment.qmd
add new section based on reviewer comment

The present work is designed to analyze time-series data of relative abundance
of multiple microbial lineages uniquely identified by a DNA barcode
[@kinsler2020; @ascensao2023]. In these competition assays, an ensemble of
genotypes is pooled together with an unlabeled reference strain that, initially,
represents the vast majority ($\geq 90\%$) of the cells in the culture (see
schematic in @fig-01(A)). Furthermore, a fraction of labeled genotypes
equivalent to the unlabeled reference strain---hereafter defined as *neutral*
lineages---are spiked in at a relatively high abundance ($\approx 3-5\%$). The
rest of the culture is left for the ensemble of genotypes of interest.

>>>>

The present work is designed to analyze time-series data of relative abundance
of multiple microbial lineages uniquely identified by a DNA barcode
[@kinsler2020; @ascensao2023]. In these competition assays, an ensemble of
genotypes is pooled together with an unlabeled reference strain that, initially,
represents the vast majority ($\geq 90\%$) of the cells in the culture (see
schematic in @fig-01(A)). Furthermore, a fraction of labeled genotypes
equivalent to the unlabeled reference strain---hereafter defined as *neutral*
lineages---are spiked in at a relatively high abundance ($\approx 3-5\%$). The
rest of the culture is left for the ensemble of genotypes of interest.
<span style="color: #912E27;">
This experimental design in which the barcodes of interest represent a small
fraction of the culture serves three main purposes: First, the model used to
infer the relative fitness differences between genotypes is valid in the regime
where the genotype frequency is significantly smaller than one (see
@sec-fitness_model). Second, potential problems with frequency-dependent
selection are minimized as long as the frequency of each genotype remains small.
Third, having a single genotype dominating the culture standardizes the
environment experienced by all genotypes. This is because variations in the
chemistry of the environment are effectively dictated by the dominating
genotype, allowing for reproducible growth-dilution cycles.
</span>
